Statistical Inference of Transport Mechanisms and Long Time Scale Behavior from Time Series of Solute Trajectories in Nanostructured Membranes

### 2020-08-11 Ben_Manuscripts/hdphmm/response_to_reviewers.tex

    %BJC2: 
    We agree that a full comparison of our method to other methods for computing long timescale 
    diffusion such as TEKMC would be very interesting and instructive but feel it would be better 
    suited as a separate study where we could evaluate a number of approaches, including perhaps
    recent approaches we have developed within our research group.~\cite{coscia_capturing_2020}
    %BJC2: added to make the point about our system being more complicated than those studied.
    It would be illuminating to test these methods on systems such as ours where there is
    a rich diversity of intermolecular interactions that work to immobilize solutes.
    Such a large-scale comparison would take more space and time than would be appropriate for 
    the current paper. Although methods such as TEKMC seems a fairly straightforward approach to
    implement, it is not immediately clear how we could optimally adapt the methodology to a 
    monoclinic unit cell with two dimensional order without in-depth testing.
    % the grids that divide the homogeneous cubic unit cells of the referenced TEKMC studies to 
    % our system which uses a monoclinic unit cell and has 2 dimensional order. 
    %BJC2: maybe push back on the CS-DBP here? 
    Regarding the CS-DBP model, we are not convinced it would be suitable for our goals because
    it relies on defining bound states and modeling hops between them. There are numerous potential
    locations within our LLC system where solutes can become bound in some way. The HDP-AR-HMM
    circumvents the need for relatively complicated methods used to identify binding sites 
    by automatically identifying distinct dynamics characteristic of a class of bound states.

    As stated by the reviewer, there are two goals of our work: to obtain long timescale MSDs and
    to understand mechanisms leading to diffusive behavior. TEKMC may be a suitable alternative 
    approach to obtaining our first goal, but it is not clear that it could give the same level
    of mechanistic insight that the HDP-AR-HMM offers. Although the work by Hanson et al.
    (J. Phys Chem. B, 2012, 116, 95-103) is able to provide physical insight into the effect of
    nanoparticles on diffusion of gaseous penetrants in a polystyrene polymer matrix, it does so
    by testing hypotheses motivated by observations of the diffusion constant (predicted by TEKMC)
    in response to different temperatures and nanoparticle concentrations. This parallels well
    with the type of analysis we performed in our initial LLC transport work (J. Phys Chem. B, 2019, 123, 6314-6330)
    where we used the mean squared displacement and the qualitative characteristics of the solute
    time series in order to hypothesize transport mechanisms. In this work we "re-discover" these
    mechanisms with minimal human intervention and characterize them in detail using the 
    HDP-AR-HMM while \textit{simultaneously} providing a method to estimate diffusion constants.

@@

%MRS: additional thoughts:

%1. Dynamic Bond Percolation theory (both the original version and the
%chemically specific one) an and TEKMC are based on spatial clustering
%and coarse-graining. Our approach is based on clustering and
%coarse-graining the dynamics.  Can be more powerful, doesn't assume
%(or minimally assumes - in some cases we assume pore vs. non-pore)
%which configurations are important, which matters when there are a
%diversity of different types of bound states, and diversity of
%dynamics (in the pore, associated with the head groups, etc).

@@

%BJC2: I believe I had orginally addressed this briefly as well. Feel free to expand if you feel there is more to be said
%3. Configurational behavior is more complex with the nanostructures,
%because one cannot simply divide the structure into cubes
%isotropically.  It may be easier to use our approach with more
%diverse configurations, because we don't need to worry about
%partitioning correctly.  This does push some of the difficulty into
%choosing the coordinate system for the dynamics, but we have a lot
%more freedom about choosing coordinate systems for the model than one
%would using a uniform (or even nonuniform) partitioning of the
%system. Yes, there are difficult choices to make but the model has
%the freedom to make these choices.

%4. One reason our simulations are longer because the timescales of
%events observed generally are longer, and the number of solutes we
%can put into the simulation is smaller (since the total pore volume
%is smaller than it would be in an amorphous polymer).

%5. We could probably construct our model it with a larger number of
%more shorter simulations as well. Rare events require long timescale
%sisulations.  However, the equilibration times for these simulations
%(since we build from geometry, not from amorphous melts) are such
%that we probably get more efficient use of simulations by running
%fewer, longer simulations.

% BJC2: assuming this comment about CS-DBP?
    % MRS2: sorry, this was about OUR method, that it's definitely different because
    % the clustering is not confiugurations based.
%6. So it's not clear that this method is better than DBP or TEKMC
% (especially since the results are a big ambiguous); one could define
% bound states, and look at hops betwen then.  But it's definitely
% DIFFERENT, it's not really the same thing at all, and we should push
% back that it is similar.  We can argue that our approach is more
% suited to our systems than the two configuration-based approaches -
% one COULD try to use those approaches and they could be made to
% work, but it's not a trivial thing to do, and we wanted to try
% something where the data informed our model more completely and we
% wouldn't have to try those modifications.

@@

    Finally, we have added a figure with a picture of the system which we are studying 
    (see~\ref{fig:membrane_side_profile}).
    \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{membrane_side_profile.pdf}
    \caption{When combined with 10 wt\% water, the LLC monomer studied in this work forms the 
    H\textsubscript{II} phase which is characterized by hexagonally packed columns with 
    hydrophilic pores. In the figure, white and red water molecules mark the membrane pores.
    The monomer fills the rest of the space. Note that water molecules are present in the 
    tails, but they are hidden for the sake of clarity. }\label{fig:membrane_side_profile}
    \end{figure}

    %r1_comment4
    \item \textit{The authors should more clearly note that their discussion largely applies only 
    to very low water content LLC regimes (such as the 10\% LLC studied). For higher water content 
    LLCs (e.g. $>$ 15-20\%) it is straightforward to obtain the diffusion limit from standard MD 
    simulations. While the authors are surely well aware of this, the reader may not be, and it is
    important to clearly distinguish the regimes.}

    Author reply: We thank the reviewer for making this point. 
    %BJC2: modded based on MRS suggestion    
    In systems with less obstructions to motion, such as LLCs with a higher water content, most 
    solutes and water will be relatively unimpeded, and diffusion constants can be obtained by 
    standard techniques.    
    However, we would still argue that our method would still be very useful for studying
    solute dynamics in these higher water content systems 
    %BJC2: incorporated MRS suggestion with small modifications    
    by allowing one to study the dynamics of solutes that have longer residence times and interact
    with moieties in a complex medium. %portions of the system away from the pore center.
    %Interactions characterized by different dynamical behavior even on short time scales can
    %still be highly informative. 
    This is the basis for the point we make in the introduction:``. . . direct calculations of diffusion 
    constants may obscure the molecular mechanisms controlling transport". Therefore we would
    prefer not to limit the range of application of this method. 
    %BJC2: added    
    However, we do agree that it would be helpful to the reader to clarify that the level of modeling 
    which we present in this work is certainly not merited in cases where they are only interested in
    diffusion constants. Therefore, we have modified a paragraph in the introduction in order to
    encourage readers to pursue simpler approaches if the system allows it.

@@

    %r1_comment8
    \item \textit{Figure captions need to be improved. Wording such as “we plotted...”, “we 
    calculated...”, “We also analyzed...” should be omitted in figure captions.}

    %BJC2: okay, modified as suggested
    Author reply: We agree with the reviewer that the type of wording described should not be
    included in captions as their primary purpose should be to explain the conclusions drawn
    by the figure rather the process of generating the figure. However, we believe that the 
    only real offender among our captions was Figure 8. Therefore we have modified the language
    of that figure:

    Original caption:
    \begin{quote}
    (a) We analyzed hydrogen bonds donated by methanol to the highlighted oxygen
    atoms in three regions of the monomer: the carboxylate groups (blue), the ether linkages
    (green) and the tails (red). (b) For each predominant state (states 1--5 described in 
    Figure~\ref{M-fig:prevalence}), we plotted the fraction of the total time in that state 
    spent hydrogen bonding to various regions of the monomer, color coded to match (a).
    (c) For each type of hydrogen bond in each predominant state, we calculated the 95th percentile
    of hydrogen bond lifetimes. (d) We also analyzed the frequency and lifetimes of solute 
    association with sodium ions. Methanol associates with sodium less than 10\% of the total 
    time it occupies any of the predominant states.
    \end{quote}

    Modified caption:
    \begin{quote}
    (a) Hydrogen bonds can be donated by methanol to the highlighted oxygen atoms in three 
    regions of the monomer: the carboxylate groups (blue), the ether linkages (green) and the
    tails (red). (b) For each predominant state (states 1--5 described in Figure~\ref{M-fig:prevalence}),
    the fraction of the total time in that state spent hydrogen bonding to various regions
    of the monomer, color coded to match (a). (c) The 95th percentile of hydrogen bond lifetimes for
    each type of hydrogen bond in each predominant state, the 95th percentile of hydrogen bond lifetimes
    (d) The frequency and lifetimes of solute association with sodium ions. Methanol associates with sodium
    less than 10\% of the total time it occupies any of the predominant states.    
    \end{quote}

### 2020-08-11 Ben_Manuscripts/hdphmm/Draft.tex
Soem comments.

  In consideration of our ultimate goal, which is to use high-throughput simulations to 
  facilitate our understanding of LLC membrane design for solute-specific separations, the size 
  of our system and length of the accompanying molecular simulations appears to be somewhat 
  impractical. Dependent on one's goals, it may not be necessary to run such long simulations in order 
  to extract valuable mechanistic insight. We originally uncovered the mechanisms discussed
  in this study using trajectories a fraction of the length here.\cite{coscia_chemically_2019} A 
  strong mechanistic understanding of solute transport with respect to the chemical features of a 
  given LLC monomer may be enough for experimentalists to eliminate entire subsets of potential
  design changes. However, it may still be valuable to consider how one might make these simulations
  amenable to high-throughput studies. We would like to emphasize that our large, fully atomistic 
  models will be extremely valuable as a benchmark system used to explore less expensive models. 
  For example, we could explore systems built with fewer LLC monomers by making the membrane pores
  shorter. We could also build systems with higher concentrations of solutes in order to gather more 
  dynamical data on shorter timescales. Using what we have learned about the radial distribution
  functions of these solutes in this and our previous work~\cite{coscia_chemically_2019}, it may be possible to 
  learn how to create initial configurations with solutes initially distributed in a way that
  is close to their equilibrium distribution, allowing solutes to thoroughly explore the 
  accessible membrane structural space more quickly and reducing the amount of unequilibrated
  data that must be discarded. Alternatively, we could use grand canonical Monte Carlo simulations
  to load solutes into the membrane~\cite{snurr_prediction_1993}. Finally, it may be possible 
  to speed up our simulations by reducing the number of atoms using united-atom representations
  of the liquid crystal monomers~\cite{horta_gromos-compatible_2016} or by using hydrogen mass 
  repartitioning to allow us to increase the MD time step.~\cite{hopkins_long-time-step_2015}

>>>>

  Given our ultimate goal, to use high-throughput simulations to 
  facilitate our understanding of LLC membrane design for solute-specific separations, the size 
  of our system and length of the accompanying molecular simulations appears may be somewhat 
  impractical. In many situations, it may not be necessary to run such long simulations in order 
  to extract valuable mechanistic insight. We originally uncovered the mechanisms discussed
  in this study using trajectories a fraction of the length here~\cite{coscia_chemically_2019}. A 
  strong mechanistic understanding of solute transport with respect to the chemical features of a 
  given LLC monomer may be enough for experimentalists to eliminate entire subsets of potential
  design changes. However, it is still valuable to consider how one might make these simulations
  amenable to high-throughput studies. We emphasize that our large, fully atomistic 
  models will be extremely valuable as a benchmark system used to explore less expensive models. 
  For example, we could explore systems built with fewer LLC monomers by making the membrane pores
  shorter. We could also build systems with higher concentrations of solutes in order to gather more 
  dynamical data on shorter timescales. Using what we have learned about the radial distribution
  functions of these solutes in this and our previous work~\cite{coscia_chemically_2019}, it may be possible to 
  learn how to create initial configurations with solutes initially distributed in a way that
  is close to their equilibrium distribution, allowing solutes to thoroughly explore the 
  accessible membrane structural space more quickly and reducing the amount of unequilibrated
  data that must be discarded. Alternatively, we could use grand canonical Monte Carlo simulations
  to load solutes into the membrane~\cite{snurr_prediction_1993}.
  If simulation resources are available, the simulations could be split up into larger numbers of shorter simulations.
  Finally, it may be possible 
  to speed up our simulations
  %MRS2: united atom would screw up the kinetics too much. 
  %by reducing the number of atoms using united-atom representations
  %of the liquid crystal monomers~\cite{horta_gromos-compatible_2016} or
  by using hydrogen mass 
  repartitioning to allow us to increase the MD time step.~\cite{hopkins_long-time-step_2015}
